---
title: "Exercise 3"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading previous data
```{r}
library(arrow)

# Loading previous data
applications <- arrow::read_feather("/Users/brettkwan/Desktop/McGill/People Analytics/app_data_starter.feather")

```

## Creating 85% subsample and 15% holdout data

```{r}
library(dplyr)

# Create the subsample (85%)
subsample <- applications %>% 
  sample_frac(0.85, replace = FALSE)

# Create the holdout data (15%)
holdout <- applications %>%
  anti_join(subsample)

```
## Count the number of unique examiner_id
```{r}
# Count the number of unique examiner_ids in the entire dataset
unique_examiner_ids <- applications %>%
  distinct(examiner_id) %>%
  nrow()

# Print the result
unique_examiner_ids

```



## attempt 1 define turnover
use disposal_type. but i also tried to account for retirements after the completion of their final application.
```{r}
# Calculate turnover rate based on disposal_type
turnover_rate <- applications %>%
  filter(disposal_type %in% c("ABN", "ISS")) %>%
  distinct(examiner_id) %>%
  nrow() / unique_examiner_ids * 100

# Print the turnover rate
cat("Overall turnover rate:", turnover_rate, "%\n")


```



## attempt 2 define turnover
Step 1: Determine application status

Identify the applications that have a disposal_type of "ISS" (issued) or "ABN" (abandoned). Exclude the applications with "PEND" (pending) as their disposal_type.
Step 2: Calculate turnover rate

Divide the number of applications with a disposal_type other than "PEND" by the total number of unique applications.
```{r}
library(tidyverse)

# Step 2: Determine application status
subsample <- subset(applications, disposal_type != "PEND")

# Step 3: Calculate turnover rate
total_applications <- n_distinct(applications$application_number)
non_pending_applications <- n_distinct(subsample$application_number)
turnover_rate <- non_pending_applications / total_applications

# Print the turnover rate
cat("Overall turnover rate:", turnover_rate)

```

### turnover rate between 2010 and 2015.
```{r}
library(dplyr)
library(lubridate)

# Convert earliest_date and latest_date to Date type
subsample <- subsample %>%
  mutate(earliest_date = as.Date(earliest_date),
         latest_date = as.Date(latest_date))

# Filter the subsample data for the time period between 2005 and 2015
filtered_data <- subsample %>%
  filter(earliest_date >= as.Date("2005-01-01") & latest_date <= as.Date("2015-12-31"))

# Join with first_last_days to identify unique examiner_id's working during the specified time period
working_examiners <- filtered_data %>%
  inner_join(first_last_days, by = "examiner_id") %>%
  filter(first_day <= as.Date("2015-01-01") & last_day >= as.Date("2015-01-01"))

# Calculate the turnover rate for the identified examiners
total_examiners <- n_distinct(working_examiners$examiner_id)
turnover_examiners <- sum(working_examiners$disposal_type %in% c("ISS", "ABN"))
turnover_rate <- turnover_examiners / total_examiners

# Print the turnover rate
cat("Turnover rate between 2005 and 2015:", turnover_rate)

```


##  turnover rate by race Between 2005 and 2015
```{r}
library(dplyr)
library(lubridate)

# Convert earliest_date and latest_date to Date type
subsample <- subsample %>%
  mutate(earliest_date = as.Date(earliest_date),
         latest_date = as.Date(latest_date))

# Filter the subsample data for the time period between 2005 and 2015
filtered_data <- subsample %>%
  filter(earliest_date >= as.Date("2005-01-01") & latest_date <= as.Date("2015-12-31"))

# Join with first_last_days to identify unique examiner_id's working during the specified time period
working_examiners <- filtered_data %>%
  inner_join(first_last_days, by = "examiner_id") %>%
  filter(first_day <= as.Date("2015-01-01") & last_day >= as.Date("2015-01-01"))

# Calculate the turnover rate by race for the identified examiners
turnover_rate_by_race <- working_examiners %>%
  group_by(race) %>%
  summarise(turnover_examiners = sum(disposal_type %in% c("ISS", "ABN")),
            total_examiners = n_distinct(examiner_id),
            turnover_rate = turnover_examiners / total_examiners)

# Print the turnover rate by race
print(turnover_rate_by_race)

```


## OLS Linear Regression - converting gender and race first
```{r}
subsample$race <- factor(subsample$race, levels = c('black', 'white', 'hispanic', 'asian'))

```
```{r}
subsample$gender <- factor(subsample$gender, levels = c('male', 'female'))

```
### ols linear regression

library(dplyr)
library(lubridate)

# Convert earliest_date and latest_date to Date type
subsample <- subsample %>%
  mutate(earliest_date = as.Date(earliest_date),
         latest_date = as.Date(latest_date))

# Filter the subsample data for the time period between 2005 and 2015
filtered_data <- subsample %>%
  filter(earliest_date >= as.Date("2005-01-01") & latest_date <= as.Date("2015-12-31"))

# Ensure 'race' contains the correct categories and is encoded as a factor variable
filtered_data$race <- factor(filtered_data$race, levels = c('black', 'white', 'hispanic', 'asian'))

# Ensure 'gender' contains the categories 'male' and 'female' and is encoded as a factor variable
filtered_data$gender <- factor(filtered_data$gender, levels = c('male', 'female'))

# Perform OLS linear regression
model <- lm(turnover_rate ~ tenure_days + race + gender, data = filtered_data)

# Make predictions on the subsample dataset
subsample_predictions <- predict(model, newdata = filtered_data)

# Print the predicted turnover values
print(subsample_predictions)


```



## Oberservations
Very difficult to find ways to define turnover. So many different variables to take into consideration. gradually, learned disposal_type would play a factor. Neeeded to understand the implications of "ISS" "PEND" and "ABN". Tenure_Days and timing of the tests would be needed as well. Tried different methods (specific time period, tenure_days, and different definitions of turnover.)
