---
title: "Exercise 5 BK"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Loading previous data
```{r }
library(tidyverse)
library(lubridate)
library(dplyr)
library(arrow)
applications <- read_feather("/Users/brettkwan/Desktop/McGill/People Analytics/app_data_starter.feather")
```

## Gender Distribution by TC
```{r gender distribution by TC}
# Assuming your dataset is called 'applications' 

# Filter the dataset to include only the relevant columns ('tc' and 'gender')
filtered_data <- applications[, c("tc", "gender")]

# Group the data by 'tc' and 'gender', and calculate the count of each gender within each group
gender_distribution <- table(filtered_data$tc, filtered_data$gender)

# Print the gender distribution by 'tc'
print(gender_distribution)

```
## Show as Ratio and Histogram
```{r}
# Assuming your dataset is called 'applications' and you have loaded the 'ggplot2' package

# Filter the dataset to include only the relevant columns ('tc' and 'gender')
filtered_data <- applications[, c("tc", "gender")]

# Calculate the gender distribution as a ratio
gender_distribution <- prop.table(table(filtered_data$tc, filtered_data$gender), margin = 1)

# Print the gender distribution as a ratio
print(gender_distribution)

# Convert the gender distribution to a data frame
gender_df <- as.data.frame(gender_distribution)

# Rename the columns for clarity
colnames(gender_df) <- c("tc", "gender", "ratio")

# Plot the stacked histogram
ggplot(gender_df, aes(x = tc, y = ratio, fill = gender)) +
  geom_bar(stat = "identity") +
  labs(x = "tc", y = "Gender Ratio", title = "Gender Distribution by tc") +
  scale_fill_manual(values = c("pink", "blue"))  # Customize the colors for each gender


```

## TC Gender Average
```{r,}
tc_gender_avg <- applications %>%
  group_by(tc, gender) %>%
  filter(gender != "n/a") %>%
  summarize(average = n(), .groups = "drop") %>%
  group_by(tc) %>%
  mutate(percentage = average/sum(average) * 100, .groups = "drop")

print(tc_gender_avg)

```

## Gender distribution by TC
```{r}
distinct_examiner_ids <- applications %>%
  filter(!is.na(gender)) %>%
  group_by(tc, gender) %>%
  filter(!is.na(examiner_id)) %>%
  summarize(distinct_count = n_distinct(examiner_id), .groups = "drop")

print(distinct_examiner_ids)

```
I remembered the example you gave in class about classroom composition perspection (90/10 and 50/50) but was confused as it how to get there.

1600 - 456 Female and 497 Male Total 953
456 Squared + 497 Squared / 953 squared = 50%
1700 - 503 Female and 859 Male Total 1362  = 53%
2100 - 346 Female and 1325 Male Total 1671  = 67%
2400 - 238 Female and 897 male. Total 1135  = 67%


## TC Gender Segregation 
```{r}
tc_gender_segregation <- tc_gender_avg %>%
  group_by(tc) %>%
  summarize(segregation_index = sum(percentage^2)/10000)
print(tc_gender_segregation)
```

## Visualize the gender representation and segregation using a bar plot.
```{r}
ggplot(tc_gender_avg, aes(x = tc, y = percentage, fill = gender)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(x = "Technical Center", y = "Percentage", fill = "Gender") +
  ggtitle("Gender Representation by Technical Center") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

##Visualize the gender segregation index using a bar plot. 
We will plot the segregation index for each TC, where higher values indicate higher segregation.
```{r}
ggplot(tc_gender_segregation, aes(x = tc, y = segregation_index)) +
  geom_bar(stat = "identity") +
  labs(x = "Technical Center", y = "Segregation Index") +
  ggtitle("Gender Segregation by Technical Center") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Examiner Art Unit average gender representation
```{r}
art_unit_gender_avg <- applications %>%
  drop_na(examiner_id) %>%
  group_by(examiner_art_unit, gender, .drop = TRUE) %>%
  summarize(average = n(), .groups = "drop") %>%
  group_by(examiner_art_unit, .drop = TRUE) %>%
  mutate(percentage = average/sum(average) * 100, .groups = "drop")
print(art_unit_gender_avg)
```


## Examiner Art Gender Segregation
```{r}
art_unit_gender_segregation <- art_unit_gender_avg %>%
  group_by(examiner_art_unit, .drop = TRUE) %>%
  summarize(segregation_index = sum(percentage^2)/10000)
```

## Visualize gender segregation index 
```{r}
ggplot(art_unit_gender_segregation, aes(x = examiner_art_unit, y = segregation_index)) +
  geom_bar(stat = "identity") +
  labs(x = "Examiner Art Unit", y = "Segregation Index") +
  ggtitle("Gender Segregation by Examiner Art Unit") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Identifying top 25 AU scoring highest on segregation index
```{r}
top_25_segregated_units <- art_unit_gender_segregation %>%
  arrange(desc(segregation_index)) %>%
  head(25)
print(top_25_segregated_units)
```
```{r}
top_25_least_segregated_units <- art_unit_gender_segregation %>%
  arrange(segregation_index) %>%
  head(25)
print(top_25_least_segregated_units)

```

## Art Unit Gender Distribution to examine differences in individual perceptions
```{r}
distinct_examiner_ids <- applications %>%
  filter(!is.na(gender)) %>%
  filter(examiner_art_unit %in% sample(unique(examiner_art_unit), 5)) %>%
  group_by(examiner_art_unit, gender) %>%
  filter(!is.na(examiner_id)) %>%
  summarize(count = n(), .groups = "drop") %>%
  group_by(examiner_art_unit) %>%
  mutate(percentage = count / sum(count) * 100)

print(distinct_examiner_ids)

```
         Versus Individ. Perception
1628        50.6%
1726        53.2%
1791        54.3%
2169        54.3%
2177        56.0
## Observations

###TC Gender Average
I didn't do distinct examinder id as I was concerned about examiners working at different TCs over the time span. Equality appears greatest in 1600, and gets worse as you ascend TC. As noted in earlier exercises, it is notably worse in 2100 and 2400.


### Segregation Index
I used something called a segregation index, which calculates the proportion of one group's population that would need to move to achieve an even distribution across all geographic areas. A value of 0 indicates complete integration, while a value of 1 indicates complete segregation.1600 was lowest with about .3499, while the others were between .41 and .45. This once again shows that 1600 is more equal. This is likely to it being a more established TC, and has been able to undergo more gradual evolutions, in line with societal change, than the other 'tc" which are newer, and may take time to catch up to societal averages.

I used this index to identify the top 25 and bottom 25 scoring Art Units according to the Segregation Index. The results match the TC gender average, inso far as the top 25 integrated art units are concentrated in 1600, while TC 2100 dominates the most most segregated.  There are some outliers, notably 2171, 2168,2179, 2176 in the top 10 of most integrated.
They all belong to either 2140/2170 Graphical User Interface and Document Processing or 2150/2160 Data Bases & File Management. Additionally 4/5 of these Art Units are lead either by someone other than a white male.  This diverse leadership could influence the gender distribution of the art unit.

Overall, Ive noticed that gender segregation/gender equality is most pronounced at the art unit level. This is evidenced by the appearance of art units in the 1600s appearing in the top 25 least integrated, and the appearance of 2100 art units in the top 25 most integrated. It seems like when baselines are set, it takes time to move towards equality, and will be done gradually. it will increase at a fast pace when there is more equal distribution. Other factors likely have to do with gender-based leadership.

###Individual Perception of Gender 
Looking at gender distribution from individual perception, the differences in gender seem to be amplified in 1600 and 1700, but downplayed in 2100 and 2400. I am not so familiar with this concept so will ask for further explanation in class.