---
title: "Exercise 2 Revised"
output: html_document
date: "2023-05-19"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(arrow)
```

## Load data

Load the following data:
  + applications from `app_data_sample.parquet`
  + edges from `edges_sample.csv`

```{r load-data}
# change to your own path!
data_path <- "/Users/brettkwan/Desktop/McGill/People Analytics/"
applications <- read_parquet(paste0(data_path,"app_data_sample.parquet"))

applications
```

## Get gender for examiners

We'll get gender based on the first name of the examiner, which is recorded in the field `examiner_name_first`. We'll use library `gender` for that, relying on a modified version of their own [example](https://cran.r-project.org/web/packages/gender/vignettes/predicting-gender.html).

Note that there are over 2 million records in the applications table -- that's because there are many records for each examiner, as many as the number of applications that examiner worked on during this time frame. Our first step therefore is to get all *unique* names in a separate list `examiner_names`. We will then guess gender for each one and will join this table back to the original dataset. So, let's get names without repetition:

```{r gender-1}
library(gender)
#install_genderdata_package() # only run this line the first time you use the package, to get data for it

# get a list of first names without repetitions
examiner_names <- applications %>% 
  distinct(examiner_name_first)

examiner_names
```

Now let's use function `gender()` as shown in the example for the package to attach a gender and probability to each name and put the results into the table `examiner_names_gender`

```{r gender-2}
# get a table of names and gender
examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )

examiner_names_gender
```

Finally, let's join that table back to our original applications data and discard the temporary tables we have just created to reduce clutter in our environment.

```{r gender-3}
# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)

# joining gender back to the dataset
applications <- applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")

# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()

```


## Guess the examiner's race

We'll now use package `wru` to estimate likely race of an examiner. Just like with gender, we'll get a list of unique names first, only now we are using surnames.

```{r race-1}
library(wru)

examiner_surnames <- applications %>% 
  select(surname = examiner_name_last) %>% 
  distinct()

examiner_surnames
```
We'll follow the instructions for the package outlined here [https://github.com/kosukeimai/wru](https://github.com/kosukeimai/wru).

```{r race-2}
examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()

examiner_race
```

As you can see, we get probabilities across five broad US Census categories: white, black, Hispanic, Asian and other. (Some of you may correctly point out that Hispanic is not a race category in the US Census, but these are the limitations of this package.)

Our final step here is to pick the race category that has the highest probability for each last name and then join the table back to the main applications table. See this example for comparing values across columns: [https://www.tidyverse.org/blog/2020/04/dplyr-1-0-0-rowwise/](https://www.tidyverse.org/blog/2020/04/dplyr-1-0-0-rowwise/). And this one for `case_when()` function: [https://dplyr.tidyverse.org/reference/case_when.html](https://dplyr.tidyverse.org/reference/case_when.html).

```{r race-3}
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))

examiner_race
```

Let's join the data back to the applications table.

```{r race-4}
# removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)

applications <- applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))

rm(examiner_race)
rm(examiner_surnames)
gc()
```


## Examiner's tenure 

To figure out the timespan for which we observe each examiner in the applications data, let's find the first and the last observed date for each examiner. We'll first get examiner IDs and application dates in a separate table, for ease of manipulation. We'll keep examiner ID (the field `examiner_id`), and earliest and latest dates for each application (`filing_date` and `appl_status_date` respectively). We'll use functions in package `lubridate` to work with date and time values.

```{r tenure-1}
library(lubridate) # to work with dates

examiner_dates <- applications %>% 
  select(examiner_id, filing_date, appl_status_date) 

examiner_dates
```

The dates look inconsistent in terms of formatting. Let's make them consistent. We'll create new variables `start_date` and `end_date`.

```{r tenure-2}
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))
```

Let's now identify the earliest and the latest date for each examiner and calculate the difference in days, which is their tenure in the organization.

```{r tenure-3}
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    ) %>% 
  filter(year(latest_date)<2018)

examiner_dates
```

Joining back to the applications data.

```{r tenure-4}
applications <- applications %>% 
  left_join(examiner_dates, by = "examiner_id")

rm(examiner_dates)
gc()
```

## Count Examiners


```{r count-examiners}
library(dplyr)

applications %>%
  distinct(examiner_id) %>%
  count()
```


## Compare Gender by TCs
```{r compare-tcs by gender}

library(dplyr)
library(ggplot2)

applications %>%
  group_by(tc, gender) %>%
  filter(!is.na(gender)) %>%
  summarise(n = n_distinct(examiner_id)) %>%
  ggplot(aes(x = as.factor(tc), y = n, fill = gender)) +
  geom_col(position = "dodge") +
  ylab("Examiners")



```

## Compare TCs by Race
```{r compare tech centers}
library(dplyr)
library(ggplot2)

applications %>%
  group_by(tc, race) %>%
  #filter(!is.na(race)) %>%
  summarise(n = n_distinct(examiner_id)) %>%
  ggplot(aes(x = as.factor(tc), y = n, fill = race)) +
  geom_col(position = "dodge") +
  ylab("Examiners")

```

## Filter TC by tenure days
```{r}
library(dplyr)
library(ggplot2)

applications %>%
  group_by(tc, tenure_days) %>%
  filter(!is.na(tenure_days)) %>%
  summarise(n = n_distinct(examiner_id)) %>%
  ggplot(aes(x = as.factor(tc), y = n, fill = tenure_days)) +
  geom_col(position = "dodge") +
  ylab("Examiners")

```

## another attempt at visualization
```{r}
library(dplyr)
library(ggplot2)

applications %>%
  group_by(tc, tenure_days) %>%
  filter(!is.na(tenure_days)) %>%
  summarise(n = n_distinct(examiner_id)) %>%
  ggplot(aes(x = as.factor(tc), y = n, fill = tenure_days)) +
  geom_col(position = "dodge") +
  ylab("Examiners") +
  scale_y_continuous(breaks = seq(500, 7500, by = 500))

```


```{r}
library(dplyr)
library(ggplot2)

applications %>%
  group_by(tc, tenure_days) %>%
  filter(!is.na(tenure_days)) %>%
  summarise(n = n_distinct(examiner_id)) %>%
  ggplot(aes(x = as.factor(tc), y = n, fill = tenure_days)) +
  geom_col(position = "dodge") +
  scale_fill_gradient(low = "blue", high = "red") +
  ylab("Examiners") +
  xlab("Technology Centers")

```

## scatter plot
```{r}
library(dplyr)
library(ggplot2)

applications %>%
  filter(!is.na(tenure_days) & tenure_days >= 4500 & tenure_days <= 6500) %>%
  ggplot(aes(x = tc, y = tenure_days)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  ylab("Tenure Days") +
  xlab("Technology Centers") +
  theme_minimal()


```
## box plot
```{r}
library(dplyr)
library(ggplot2)

applications %>%
  filter(!is.na(tenure_days) & tenure_days >= 00 & tenure_days <= 6500) %>%
  ggplot(aes(x = tc, y = tenure_days)) +
  geom_violin(fill = "lightblue", color = "black") +
  ylab("Tenure Days") +
  xlab("Technology Centers") +
  theme_minimal()


```

## another attempt
```{r}
library(dplyr)
library(ggplot2)

applications %>%
  group_by(tc, race) %>%
  summarise(n = n_distinct(examiner_id), avg_tenure = mean(tenure_days)) %>%
  ggplot(aes(x = reorder(tc, avg_tenure), y = avg_tenure, fill = race)) +
  geom_col(position = "dodge") +
  ylab("Average Tenure (in days)") +
  xlab("Technology Centers") +
  ggtitle("Average Tenure of Examiners in Different Technology Centers") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("blue", "red", "green", "yellow", "orange")) # Adjust color values as per your preference

```



### Correlations


```{r}

library(dplyr)
library(ggplot2)
examiners <- applications %>%
  group_by(examiner_id) %>%
  summarise(
    tenure = first(tenure_days), 
    gender = first(gender),
    race = first(race)
    )

library(broom)

fit <- lm(tenure ~ gender + race, data = examiners)
tidy(fit)


```


## Questions
Why is black over represented in this analysis?

## write out to file, to save time in future exercises
```{r}
write_feather(applications, paste0(data_path,"app_data_starter.feather"))

```

## Open data file

I'm going to load the data I saved after the last exercise.


applications <- read_feather("/Users/brettkwan/Desktop/McGill/People Analytics/app_data_starter.feather")


### Oberservations
-	Males outnumber females by a significant margin, especially 2100 and 2400, which focus on Computer Architecture and Networks.
-	Males and females are most equal in BioTechnology. Females are similarly distributed in Chemistry as well.
-	Whites dominate every TC, followed by Asian, Black and Hispanic.
-	On Average, people stay more at Chemistry. Computers Networks has the shortest window. Could be due to it being a newer field and more tenured employees may not have the expertise to work in it.




